Cyclone: Fault Tolerance Middleware for NVM Clusters

### 2016-02-10 paper/paper.tex
More text in paper.

\section{Programming Model}
Cyclone provides users with a replicated

\section{Failure Atomicity}

>>>>

\section{Programming Model}
Cyclone provides users with a replicated and durable heap that can be
manipulated via a set of library routines. The library provides special
functions to allocate and free persistent memory as well as fat pointers that
wrap pointers to persistent memory to ensure that they do not change across
restarts - since the virtual memory ranges into which persistent memory is
mapped may change.

\begin{figure}
\begin{verbatim}

struct ll_node {
  int value;
  TOID(ll_node) next;
};

void insert_after(TOID(ll_node) prev, TOID(ll_node) new_node)
{
  D_RW(new_node)->next = D_RO(prev)->next;
  D_RW(prev)->next = new_node;
}

\end{verbatim}
\caption{A Persistent Linked List of Integers}
\label{fig:example}
\end{figure}

Figure~\ref{fig:example} shows the example of a linked list node and a function
for inserting a new node after a given one. The {\tt TOID} declaration wraps the
linked list node type to produce a fat pointer to it. A {\tt TOID(ll\_node)}
type therefore is a pointer to a linked list node stored in persistent memory.
The {\tt insert\_after} function used the {\tt D\_RO} and {\tt D\_RW} calls to
convert fat pointers into directly accessible read-only and read-write pointers
respectively. Other than the mechanics of using fat pointers the code entirely
resembles that which a programmer might write for a standard linked list in
DRAM.

\section{Failure Atomicity}
Programmers typically find it hard to reason about failures when writing
code. This is a problem when directly manipulating data structures in persistent
memory. Depending on the specifics of the system architecture, updates to
persistent memory can be persisted out of order thereby leaving  the data
structure in an inconsistent state. To illustrate this consider a failure during
execution of the code to add a node to a linked list in
Figure~\ref{fig:example}. If either of the two updates to the next fields of the
previous or new node fail to be persisted, the linked list is left in a
disconnected state with one or more connected nodes inaccessible.

Cyclone provides tools for programmers to write code that is failure
atomic. Portions of the code are delimited as transactions that are either
completely executed in terms of its effects on persistent memory or not executed
at all. Under the hood, Cyclone tracks all updates to persistent memory in a
persistent undo log - the log entries are persisted before the update to the
actual memory is allowed. On a failure, the undo log is used to roll back
incomplete transactions. A transaction is considered complete when it completes
execution and the updated objects are flushed (automatically by Cyclone) to
persistent memory. The log entries for the transaction are then immediately
deleted. The only burden placed on a programmer for failure atomicity is that
they must inform the Cyclone runtime about objects that are
updated. Figure~\ref{fig:example_fatomic} shows how to convert the function for
adding a node to a linked list into a failure atomic one. The only additions are
the transaction delimiters and the {\tt TX\_ADD} calls to mark the objects being
updated \emph{before} the actual updates to them.

\begin{figure}
\begin{verbatim}

struct ll_node {
  int value;
  TOID(ll_node) next;
};

void insert_after(TOID(ll_node) prev, TOID(ll_node) new_node)
{
  TX_BEGIN {
    TX_ADD(new_node);
    D_RW(new_node)->next = D_RO(prev)->next;
    TX_ADD(prev);
    D_RW(prev)->next = new_node;
  } TX_END
}

\end{verbatim}
\caption{A Persistent Linked List of Integers}
\label{fig:example}
\end{figure}
